---
title: "Data606 Homework 1"
author: "Joseph Simone"
date: "September 1, 2019"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
options(tinytex.verbose = TRUE)
```
##  Smoking Habits of UK Residents (1.10, p.20)


```{r }
smokinghabits <- read.csv("https://raw.githubusercontent.com/jbryer/DATA606Fall2019/master/course_data/os3_data/Ch%201%20Exercise%20Data/smoking.csv")
```

A. What does each row of the data matrix represent ?



```{r}
head(smokinghabits)
```
Solution: Each row of the matrix represents one observation. In this dataframe, this is a single resident of  the UK. 

B. How many participants were included in the survey ? 

```{r}
nrow(smokinghabits)
```

C. Indicate whether each variable in the study is numerical or categorical. If numerical, identify as continuous or discrete. If categorical, indicate if the variable is ordinal.

```{r}
str(smokinghabits)
```
 
Solution: 1691 particants in this survey. 


##  Cheater, scope of inference (1.14, p.29)


Exercise 1.5 introduces a study where researchers studying the rela-
tionship between honesty, age, and self-control conducted an experiment on 160 children between the ages
of 5 and 15. The researchers asked each child to toss a fair coin in private and to record the outcome (white
or black) on a paper sheet, and said they would only reward children who report white. Half the students
were explicitly told not to cheat and the others were not given any explicit instructions. Differences were
observed in the cheating rates in the instruction and no instruction groups, as well as some differences across
children's characteristics within each group.



(a) Identify the population of interest and the sample in this study.

Solution: The population of interest are children between the ages of 5 to 15. 


(b) Comment on whether or not the results of the study can be generalized to the population, and if the
findings of the study can be used to establish causal relationships.


Solution: This is an experiment, the findings of this experiment are used to establish causal relationships. 


##  Reading the Paper (1.28, p.31)


Below are excerpts from two articles published in the NY Times:
(a) An article titled Risks: Smokers Found More Prone to Dementia states the following:

"25 Researchers analyzed data from 23,123 health plan members who participated in a voluntary exam and
health behavior survey from 1978 to 1985, when they were 50-60 years old. 23 years later, about 25% of
the group had dementia, including 1,136 with Alzheimer's disease and 416 with vascular dementia. After
adjusting for other factors, the researchers concluded that pack-a-day smokers were 37% more likely than
nonsmokers to develop dementia, and the risks went up with increased smoking; 44% for one to two packs
a day; and twice the risk for more than two packs."

Based on this study, can we conclude that smoking causes dementia later in life? Explain your reasoning.

Solution: There is no treatment or control group, therefore this is an observational study. Casual relationships between variables cannot be defined in an observational study. Even though, the numbers show a correlation between smoking and dementia, we are not defining all the external factors can can be contributing to these cases. 

(b) Another article titled The School Bully Is Sleepy states the following:

"26 The University of Michigan study, collected survey data from parents on each child's sleep habits and
asked both parents and teachers to assess behavioral concerns. About a third of the students studied were
identifed by parents or teachers as having problems with disruptive behavior or bullying. The researchers
found that children who had behavioral issues and those who were identied as bullies were twice as likely
to have shown symptoms of sleep disorders."

A friend of yours who read the article says, 
"The study shows that sleep disorders lead to bullying in school children." 


Is this statement justifed? If not, how best can you describe the conclusion that can
be drawn from this study?

Solution: This is also an observational study. We cannot say for certain if the sleep disorders in these cases would lead to bullying.


##  Exercise and Mental Health (1.34, p.35)

A researcher is interested in the effects of exercise on mental health
and he proposes the following study: Use stratifed random sampling to ensure representative proportions
of 18-30, 31-40 and 41- 55 year olds from the population. Next, randomly assign half the subjects from each
age group to exercise twice a week, and instruct the rest not to exercise. Conduct a mental health exam at
the beginning and at the end of the study, and compare the results.

(a) What type of study is this?

  Solution: This is a Prospective experimental Study
            
(b) What are the treatment and control groups in this study?
     
Solution: 
      Treatment Group: exercise 2x's a week
         Control Group: no exercise


(c) Does this study make use of blocking? If so, what is the blocking variable?
    
Solution: This study does make use of blinding, the age of the groups is the blocking variable. 

(d) Does this study make use of blinding?

Solution: This study does not make use of blinding, patients and doctors are completely aware of the groups. 
    
    
(e) Comment on whether or not the results of the study can be used to establish a causal relationship
between exercise and mental health, and indicate whether or not the conclusions can be generalized to
the population at large.

Solution: Conclusions can be drawn from the population since this experiment was randomized.

(f) Suppose you are given the task of determining if this proposed study should get funding. Would you
have any reservations about the study proposal?

Solution: I would not fun this project on the basis that cluster sampling would be more benefical to the study. 


