---
title: "Data 606 - Chapter 8 - Introduction to Linear Regression"
author: "Joseph Simone"
date: "11/5/2019"
output:
  html_document:
  pdf_document: default
---

```{r setup, include=FALSE}
library(png)
library(knitr)
library(openintro)
```

### Nutrition at Starbucks, Part I. (8.22, p. 326)

```{r, include=TRUE }
imgage <- "C:/Users/jpsim/Documents/Stat & Probability for Data/starbucks.png"
include_graphics(imgage)
```

(a) Describe the relationship between number of calories and amount of carbohydrates (in grams) that
Starbucks food menu items contain.

There is positive linear relationship between calorie and carbohydrates.

(b) In this scenario, what are the explanatory and response variables?


x - axis is Calorie which is explanatory variable

y-axis is Carbohydrate which is response variable.


(c) Why might we want to fit a regression line to these data?

We would like to predict the amount of carbs based on calorie count.


(d) Do these data meet the conditions required for fitting a least squares line?

The data fit a linear plot, residuals appear nearly normal. we cannot achieve constant variability.


### Body measurements, Part I. (8.13, p. 316)

```{r, include=TRUE }
imgage <- "C:/Users/jpsim/Documents/Stat & Probability for Data/body.png"
include_graphics(imgage)
```

```{r}
data(bdims)
```

```{r}
par(mar = c(3.8, 3.8, 0.5, 0.5), las = 1, mgp = c(2.7, 0.7, 0), 
    cex.lab = 1.25, cex.axis = 1.25)
plot(bdims$hgt ~ bdims$sho.gi, 
     xlab = "Shoulder girth (cm)", ylab = "Height (cm)", 
     pch = 19, col = COL[1,2])
```

(a) Describe the relationship between shoulder girth and height.
This relationship shows that the larger ones shoulder girth the taller one might be.
 

(b) How would the relationship change if shoulder girth was measured in inches while the units of height
remained in centimeters?

This relationship would remain the same.

## Body measurements, Part III. (8.24, p. 326)

Exercise above introduces data on shoulder girth and height of a group of individuals. The mean shoulder girth is 107.20 cm with a standard deviation of 10.37 cm. The mean height is 171.14 cm with a standard deviation of 9.41 cm. The correlation between height and shoulder
girth is 0.67.

(a) Write the equation of the regression line for predicting height.

 $\hat{y} = 105.8445 + 0.6091 * `shouldergirth`$
 
(b) Interpret the slope and the intercept in this context.

For additional cm of shoulder girth, there would be additional .6091 of height.We would expect a height of .6091 , If the shoulder girth of person is Zero.


(c) Calculate R2 of the regression line for predicting height from shoulder girth, and interpret it in the
context of the application.

```{r}
B1=round(0.67 * (9.41/10.35),4)
B0=round(B1 * -107.20 + 171.14 , 4)
R=0.67
R2=R*R
R2
```

(d) A randomly selected student from your class has a shoulder girth of 100 cm. Predict the height of this
student using the model.

```{r}
R100=B0 + B1 * 100
R100
```


(e) The student from part (d) is 160 cm tall. Calculate the residual, and explain what this residual means.

```{r}
Res=160 - 166.7545
Res
```

Model Overestimated the height of the individual.

(f) A one year old has a shoulder girth of 56 cm. Would it be appropriate to use this linear model to
predict the height of this child?

This can be achieved only through Extrapolation. We are making an assumption that we can achieve linear relationship in uncharted data


## Cats, Part I. (8.26, p. 327) 

```{r, include=TRUE }
imgage <- "C:/Users/jpsim/Documents/Stat & Probability for Data/cats.png"
include_graphics(imgage)
```

(a) Write out the linear model.

$\hat{y} = -0.357 + 4.034 * `bodyweight`$

(b) Interpret the intercept.

If cat’s body weight is zero, we will expect the heart to weight -0.357 grams

(c) Interpret the slope.

For each additional kg of body weight, we can expect cat’s heaert to weigh additional 4.034 grams

(d) Interpret R2

$R2 = 64.66%$, which means 64.66% of observed data can be explained using the linear model in a 

(e) Calculate the correlation coefficient.

```{r}
B0=-0.357
B1=4.034

R2=.6466
corcof = sqrt(R2)
corcof
```

## Rate my professor. (8.44, p. 340)
```{r, include=TRUE }
imgage <- "C:/Users/jpsim/Documents/Stat & Probability for Data/ratep.png"
include_graphics(imgage)
```

(a) Given that the average standardized beauty score is -0.0883 and average teaching evaluation score
is 3.9983, calculate the slope. Alternatively, the slope may be computed using just the information
provided in the model summary table.

```{r}
B0=4.010
B1=4.13 * 0.0322
B1
```


(b) Do these data provide convincing evidence that the slope of the relationship between teaching evaluation
and beauty is positive? Explain your reasoning.

On looking at scatter plot we can just see Scaters. There is no any upward or downward trend. p is shown as zero in summary table. It can be read as accepting the null hypothesis; there is no relation between teaching evaluation and beauty.

(c) List the conditions required for linear regression and check if each one is satisfied for this model based
on the following diagnostic plots.

```{r, include=TRUE }
imgage <- "C:/Users/jpsim/Documents/Stat & Probability for Data/ratepc.png"
include_graphics(imgage)
```

```{r, include=TRUE }
imgage <- "C:/Users/jpsim/Documents/Stat & Probability for Data/ratepca.png"
include_graphics(imgage)
```

Visual inspection of scatterplot suggest residuals are randomly scattered around horizontal axis. Therefore, linearity is achieved.

Histogram skewed to the left, resulting in the possibility of some outliers. 

Residuals are normal.

In the scatterplot, the points has constant variance.

These are independant observations, and shows a minor linear trends.