---
title: "Data 606 - Final Project"
author: "Joseph Simone"
date: "12/8/2019"
output: pdf_document
---

## Linear Regression - Weather Data

```{r setup, echo=TRUE, warning=FALSE, message=FALSE}
library(RCurl)
library(rattle)
library(batchtools)
library(ggplot2)
library(caret)
library(gmodels)
library(gridExtra)
library(corrplot)
library(e1071)
library(dplyr)
library(GGally)
```

### First Data-Import 

```{r, warning=FALSE, message=FALSE}
weather_raw <- read.csv(url("https://raw.githubusercontent.com/josephsimone/DATA607/master/ww-ii-data..csv"), header = TRUE, sep = ",", dec=".", stringsAsFactors = TRUE)
```


```{r, echo=TRUE, warning=FALSE, message=FALSE}
dim(weather_raw)
```


```{r, , warning=FALSE, message=FALSE}
str(weather_raw)
```


### Cases 

To begin with, I wanted to do a Linear Regression Model using this War World II that I found on [Kaggle](https://www.kaggle.com/smid80/weatherww2), however, as you can see by the Data Types of the this Data-Set.  Since this Data-Set is from the 1940's, some of the values had 'character' values mixed in with 'decimal' valus. Nevertheless, I would like to continue to use "Weather-Related" Data. Therefore, I will for using the Sample Data-Set from ["Data Mining with Rattle and R The Art of Excavating Data for Knowledge Discover" by Graham Williams](https://link.springer.com/book/10.1007/978-1-4419-9890-3), "weather". This Data-Set comes from a weather monitoring station located in Canberra, Australia.The Bureau makes available 13 months of daily weather observations from many locations across Australia. 


At first glace, the information in this dataset includes  Rainfall, temperatures, wind speed and  humatity at different times of day.There are over twenty cases in this dataset, however, a lot of the cases and NULL. Therefore, I will be eleminating them from my dataset for analysis. The cases that I will be keeping include, the station #, , date, precipitation, wind gust speed, max, min and mean temperature, and snowfall.**


### Data collection 

Observations were drawn from Canberra Airport {station 070351}.

Some cloud observations are from automated equipment; these are somewhat different to those made by a human observer and may not appear every day.


### Type of study 

This study is an observational study, this is a collected of weather conditions recorded on each day at various weather station in Australia.


### Data Source 

[Australian Government - Bureau of Meterology](www.ncdc.noaa.gov/data-access/land-based-station-data/land-based-datasets/world-war-ii-era-data).

["Data Mining with Rattle and R The Art of Excavating Data for Knowledge Discover" by Graham Williams](https://link.springer.com/book/10.1007/978-1-4419-9890-3)

### Dependent Variable

The response vairbale that I will be trying to calculate, is going to be Rainfall. Given the Minimum Temperature can you predict the amount of Rainfal. Therefore, this is a quantitative variable.**

### Independent Variable

The two independent variables that I will using for this linear regression analysis are the Rainfall and Minimum Tempatures.

### Final Data-Import 

```{r}
dim(weather)
```

```{r}
names(weather)
```


##### Variable Distribtion:

Date: The date of observation (a date object)<br>
Location: The common name of the location of the weather station<br>
MinTemp: The minimum temperature in degrees centigrade<br>
MaxTemp: The maximum temperature in degrees centigrade<br>
Rainfall: The amount of rainfall recorded for the day in millimeters<br>
Evaporation: Class A pan evaporation (in millimeters) during 24 h<br>
Sunshine: The number of hours of bright sunshine in the day<br>
WindGustDir: The direction of the strongest wind gust in the 24 h to midnight<br>
WindGustSpeed: The speed (in kilometers per hour) of the strongest wind gust in the 24 h to midnight<br>
WindDir9am: The direction of the wind gust at 9 a.m.<br>
WindDir3pm: The direction of the wind gust at 3 p.m.<br>
WindSpeed9am: Wind speed (in kilometers per hour) averaged over 10 min before 9 a.m.<br>
WindSpeed3pm: Wind speed (in kilometers per hour) averaged over 10 min before 3 p.m.<br>
Humidity9am: Relative humidity (in percent) at 9 a.m.<br>
Humidity3pm: Relative humidity (in percent) at 3 p.m.<br>
Pressure9am: Atmospheric pressure (hpa) reduced to mean sea level at 9 a.m.<br>
Pressure3pm: Atmospheric pressure (hpa) reduced to mean sea level at 3 p.m.<br>
Cloud9am: Fraction of sky obscured by cloud at 9 a.m. This is measured in ”oktas,” which are a unit of eighths. It records how many eighths of the sky are obscured by cloud. A 0 measure indicates completely clear sky, while an 8 indicates that it is completely overcast<br>
Cloud3pm: Fraction of sky obscured by cloud at 3 p.m; see Cloud9am for a description of the values<br>
Temp9am: Temperature (degrees C) at 9 a.m.<br>
Temp3pm: Temperature (degrees C) at 3 p.m.
RainToday: Integer 1 if precipitation (in millimeters) in the 24 h to 9 a.m. exceeds 1 mm, otherwise 0<br>
RISK_MM: The continuous target variable; the amount of rain recorded during the next day<br>
RainTomorrow: The binary target variable whether it rains or not during the next day<br>

```{r}
write.csv(weather, file = "weather.csv")
```


```{r}
set.seed(123)
wd <- read.csv(url("https://raw.githubusercontent.com/josephsimone/DATA607/master/weather.csv"), header = TRUE, sep = ",", stringsAsFactors = TRUE)

dim(wd)
```


```{r}
wd <- wd[complete.cases(wd),]
```


```{r}

wd <- subset(wd, select = -c(Date, Location, RISK_MM, RainTomorrow, RainToday, WindDir9am, WindDir3pm))
wd <- wd%>%
   select(Rainfall, everything())

```



### Reserach Question

##### Given Minimum Temperature of a Day, can one predicted the amount of RainFall that it will rain the following day ?

$H_\theta:$ Given Minimum Temperature, can you predict the amount of Rainfall for that day ?
$H_A:$ One cannot predict the amount of Rainfall in a day, given the Minimum Temperature.
### Scatter Plot

```{r, warning=FALSE, message=FALSE}
scatter.smooth(x=wd$Rainfall, y=wd$MinTemp, main="")
```


### BoxPlot
```{r}
par(mfrow=c(1, 2))

boxplot(wd$Rainfall, main="Rainfall", sub=paste("Outlier rows: ", boxplot.stats(wd$Rainfall)$out))
boxplot(wd$MinTemp, main="MinTemp", sub=paste("Outlier rows: ", boxplot.stats(wd$MinTemp)$out))
```


### Density PLot


```{r}
par(mfrow=c(1, 2))

plot(density(wd$Rainfall), main="Density Plot: Rainfall", ylab="Frequency", sub=paste("Skewness:", round(e1071::skewness(wd$Rainfall), 2)))
plot(density(wd$MinTemp), main="Density Plot: MinTemp", ylab="Frequency", sub=paste("Skewness:", round(e1071::skewness(wd$MinTemp), 2)))
polygon(density(wd$MinTemp), col="blue")

```

### Correlation between Rainfall and MinTemp

```{r}
cor(wd$Rainfall, wd$MinTemp)
```

```{r}
ggpairs(data=wd, columns=1:2, title="Rainfall vs MinTemp")
```


### Linear Model

```{r}
linearMod <- lm(Rainfall~MinTemp, data=wd)
print(linearMod)
```

```{r}
summary(linearMod)
```


### Pvalue 
```{r}
lmp <- function (modelobject) {
    if (class(modelobject) != "lm") stop("Not an object of class 'lm' ")
    f <- summary(modelobject)$fstatistic
    p <- pf(f[1],f[2],f[3],lower.tail=F)
    attributes(p) <- NULL
    return(p)
}
lmp(linearMod)
```

### R-Squared

```{r}
summary(linearMod)$r.squared
```


### Conclusion
The correlation between the Rainfall and Minimum Temperature is very weak. In addition the $R^2$ and $P$ value for this Linear Model are very low. 

Thereofore we reject the null hypothesis, in favor of the alternative hypothesis, you cannot predict the amount of Rainfall given the Minimum Temperature.

Weather patterns are known to be none-linear.  

In the future,I would like to include more variables than the Maximum Temperature to predict the null hypothesis.