---
title: "Data 606 Assignment 3"
author: "Joseph Simone"
date: "9/13/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```
## Chapter 3 - Porbability 

Libraries used:
```{r}
library(knitr)
library(VennDiagram)
library(png)
library(grid)
library(gmodels)
library(openintro)
```


## Dice Roll
```{r, include=TRUE }
imgage <- "C:/Users/jpsim/Documents/Stat & Probability for Data/1.png"
include_graphics(imgage)
```

a.There is not combinations a pair of dice being rolled will result in a sum of 1.

b. 
```{r}
## Combinations: ((1,4),(2,3), (3,2), (4,1))

(4/36)
```

c. 
```{r}
#Combinations (6,6)
(1/36)
```

## Poverty and Language 
```{r, include=TRUE }
imgage <- "C:/Users/jpsim/Documents/Stat & Probability for Data/2.png"
include_graphics(imgage)
```

a.There are 4.2% of people who fall into both living below the proverty line and speaking a foreign language, therefore, the answer is no.

b.

```{r}
venn <- draw.pairwise.venn(14.6, 20.7, 4.2, c("Foreing Language", "Povery"),  scale = FALSE, fill = c("blue", "red"));
grid.draw(venn);
```

c.  10.4 percent living below the poverty line and speak English at home

d.

```{r}
Poverty_foreign <- (14.6+20.7)-4.2
Poverty_foreign
```

e.

```{r}
100 - Poverty_foreign
```


f. 

## Assortative Mating
```{r, include=TRUE }
imgage <- "C:/Users/jpsim/Documents/Stat & Probability for Data/3.png"
include_graphics(imgage)
```

```{r}
am <- data.frame(assortive.mating)
am
```


```{r}
blue = subset(am, partner_female == "blue")

x = table(am$self_male, am$partner_female)
y = as.data.frame(x)
names(y)[1] = 'Self Male'
names(y)[2] = 'Partner Female'
y
```

a.
```{r}
 ((sum(am$self_male =="blue")/nrow(am)) +
+  (sum(am$partner_female =="blue")/nrow(am)))- (sum(am$self_male =="blue" & am$partner_female=="blue")/nrow(am))
```


b. 

```{r}
 (sum(am$self_male =="blue" & am$partner_female=="blue")/nrow(am))/(sum(am$self_male=="blue")/nrow(am))
```

c. 
```{r}
(sum(am$self_male =="brown" & am$partner_female=="blue")/nrow(am))/(sum(am$self_male=="brown")/nrow(am))
```

d. The probability that a random self male from the study has a blue eyes is 56% and the probability of randomly select a partner female with blue eyes is 52.4% .Therefore, Both variables are dependent.

## Books on a Bookshelf
```{r, include=TRUE }
imgage <- "C:/Users/jpsim/Documents/Stat & Probability for Data/4.png"
include_graphics(imgage)
```


```{r}
bc <- data.frame(books)

```

```{r}
with(warpbreaks, CrossTable(bc$type, bc$format, prop.r = TRUE, prop.c = FALSE, prop.t = FALSE, prop.chisq = FALSE))
```


a.

```{r}
(sum(bc$format=="hardcover"))/sum(table(bc)) *
(sum(bc$format=="paperback" & bc$type == "fiction"))/(sum(table(bc))-1)
```


b. 

```{r}
(sum(bc$type=="fiction"))/sum(table(bc)) *
(sum(bc$format=="hardcover"))/(sum(table(bc))-1)
```


c.

```{r}
(sum(bc$type=="fiction"))/sum(table(bc)) *
(sum(bc$format=="hardcover"))/(sum(table(bc)))
```

d. Explain why this is the case. The larger the samples the smaller the difference.

## Baggage fees
```{r, include=TRUE }
```


```{r, include=TRUE }
imgage <- "C:/Users/jpsim/Documents/Stat & Probability for Data/5.png"
include_graphics(imgage)
```


a. 

```{r}
num_bags <- c('no-bags', 'one-bags', 'two-bags')
bag_fee <- c(0,25,35)
pass <- c(.54, .34, .12)
sample <- data.frame(num_bags, pass, bag_fee)
average_revenue <- (sum((sample$pass*sample$bag_fee))/sum(sample$pass))
average_revenue
```


```{r}
sqrt(0.54*(0-average_revenue)^2 + 0.34*(25-average_revenue)^2 + 0.12*(35-average_revenue)^2)
```


b.

```{r}
avg120flight <- (65*0 + 41*25 + 14 *25 + 14 * 35)
avg120flight
```

```{r}
x <- c(0, 1025,840)

sqrt(sum((x-mean(x))^2/(length(x)-1)))
```

```{r}
sd(x)
```


## Income and gender
```{r, include=TRUE }
imgage <- "C:/Users/jpsim/Documents/Stat & Probability for Data/5.png"
include_graphics(imgage)
```

a. 

```{r}
Income_range <- c("1 to $9999 or loss",
                  "10,000 to 14,999", "15, 000 to 24,999",
                  "25,000 to 34,999", "35,000 to 49,999", 
                  "50,000 to 64,999", "65,000 to 74,999",
                  "75,000 to 99,999", "100,000 or more")
total <- c(.022, .047, 0.158, 0.183, 0.212, 0.139, 0.058, 0.084, 0.097)
x <- data.frame (Income_range, total)
x
```


```{r}
barplot(total)
```


b.

```{r}
sum(x[1:5,2])
```


c.

```{r}
0.622 * 0.41
```


d. In conculsion, the assumption that was made in part C is wrong,it appears that like events are not independent.
