---
title: "Data 605 Assingment #2 "
author: "Joseph Simone"
date: "9/7/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



## Problem Set 1 

1. Proof 

```{r}
a <- matrix(1:12, ncol = 3, nrow = 4)

a
```


```{r}
at<-t(a)

aat<-a%*%at

aat
```

```{r}
ata<-at%*%a

ata
```

2. For a special type of square matrix A, Under what conditions could this be true?

```{r}
a <- matrix(c(1,2,3,2,4,5,3,5,6), ncol = 3, nrow = 3)

a
```

```{r}
at<-t(a)

att<-a%*%at

att
```


```{r}
ata<-at%*%a

ata
```


## Problem Set 2 

1. Matrix factorization is a very important problem. There are supercomputers built just to do matrix factorizations. Every second you are on an airplane, matrices are being
factorized. Radars that track flights use a technique called Kalman filtering. At the heart of Kalman Filtering is a Matrix Factorization operation. Kalman Filters are solving linear systems of equations when they track your flight using radars.

Write an R function to factorize a square matrix A into LU or LDU, whichever you prefer.

```{r}
x<-matrix(c(11,4,9,8,88,7,1,77,32,3,99,67,67,45,3,5),ncol=4,nrow=4)

x
```


```{r}
l<-function(b){
  A<-diag(nrow(b))
  for (i in 2:(nrow(b)))
    {for (j in i:(nrow(b)))
      {
      temp<-b[j,i-1]/b[i-1,i-1]
      b[j,]<-b[j,]-(b[j,i-1]/b[i-1,i-1])*b[i-1,]
      A[j,i-1]<-temp
      }
    }
    print(b)
    print(A)
    return(A%*%b)
            }

A<-l(x)
```



```{r}
A
```

